---
title: "Foltz_Correlation-analysis_2021-09-22"
output: html_notebook
---

Correlation analysis

# Packages
```{r}
library('metafor')
library('tidyverse')
library('dplyr')
library('ggplot2')
```

# Read in data and prepare meta variables
```{r}
meta.data <- read.csv("N2O ratio database 2021-09-22.csv", header=TRUE, sep=",")

# make dataframe of samples that have treatment and control
meta.data.wcont.wsd <- meta.data %>%
  filter(!is.na(N2O.ratio_cont), SD_cont > 0, SD_trt > 0)

# Use escalc function to setup meta using mean difference "MD" --> calculates yi and vi
dat <- escalc(measure = "MD", m1i = N2O.ratio_trt, m2i = N2O.ratio_cont,
              sd1i = SD_trt, sd2i = SD_cont, n1i = n, n2i = n,
              data = meta.data.wcont.wsd, append = TRUE)
```

### Begin by looking at data that have treatment and control
# Test data for normality 
```{r}
shapiro.test(dat$N2O.ratio_trt)
shapiro.test(dat$N2O.ratio_cont)
shapiro.test(dat$Soil.pH)
shapiro.test(dat$C.added)
shapiro.test(dat$N.added)
shapiro.test(dat$C.N.added)
shapiro.test(dat$Total.C)
shapiro.test(dat$Total.N)
shapiro.test(dat$C.N.total)
shapiro.test(dat$yi)
shapiro.test(dat$vi)
#hist(dat$N2O.ratio_trt,main="Histogram of cumulative N2O fluxes",xlab="N2O flux (kgN/ha)")
```
Non-normal data, so use Spearman's rank correlation test

# Correlations
# testing treatment effect (yi) vs. several variables also tested in the meta
```{r}
cor.test(dat$yi, dat$Soil.pH, method="spearman") 
cor.test(dat$yi, dat$C.added, method="spearman") 
cor.test(dat$yi, dat$N.added, method="spearman") 
cor.test(dat$yi, dat$C.N.added, method="spearman") 
cor.test(dat$yi, dat$Total.C, method="spearman") 
cor.test(dat$yi, dat$Total.N, method="spearman") 
cor.test(dat$yi, dat$C.N.total, method="spearman") 
```

# testing N2O ratio trt vs. several variables also tested in the meta
```{r}
cor.test(dat$N2O.ratio_trt, dat$Soil.pH, method="spearman") 
cor.test(dat$N2O.ratio_trt, dat$C.added, method="spearman") 
cor.test(dat$N2O.ratio_trt, dat$N.added, method="spearman") 
cor.test(dat$N2O.ratio_trt, dat$C.N.added, method="spearman") 
cor.test(dat$N2O.ratio_trt, dat$Total.C, method="spearman") 
cor.test(dat$N2O.ratio_trt, dat$Total.N, method="spearman") 
cor.test(dat$N2O.ratio_trt, dat$C.N.total, method="spearman") 
```

### Next consider data that have treatments (without excluding those without controls)
# Test data for normality 
```{r}
shapiro.test(meta.data$N2O.ratio_trt)
shapiro.test(meta.data$N2O.ratio_cont)
shapiro.test(meta.data$Soil.pH)
shapiro.test(meta.data$C.added)
shapiro.test(meta.data$N.added)
shapiro.test(meta.data$C.N.added)
shapiro.test(meta.data$Total.C)
shapiro.test(meta.data$Total.N)
shapiro.test(meta.data$C.N.total)
```

# Nonnormal data so continue using Spearmans (nonparametric)
# testing N2O ratio trt (full database) vs. several variables also tested in the meta then organizing into table
```{r}
cor1 <- cor.test(meta.data$N2O.ratio_trt, meta.data$Soil.pH, method="spearman") 
cor2 <- cor.test(meta.data$N2O.ratio_trt, meta.data$C.added, method="spearman") 
cor3 <- cor.test(meta.data$N2O.ratio_trt, meta.data$N.added, method="spearman") 
cor4 <- cor.test(meta.data$N2O.ratio_trt, meta.data$C.N.added, method="spearman") 
cor5 <- cor.test(meta.data$N2O.ratio_trt, meta.data$Total.C, method="spearman") 
cor6 <- cor.test(meta.data$N2O.ratio_trt, meta.data$Total.N, method="spearman") 
cor7 <- cor.test(meta.data$N2O.ratio_trt, meta.data$C.N.total, method="spearman") 

corr.df <- data.frame(variable=c("pH", "C added","N added", "C/N added", "Total C", "Total N", "Total C/N"), rho=1:7)

corr.df$rho[1] <- print(cor1$estimate)
corr.df$rho[2] <- print(cor2$estimate)
corr.df$rho[3] <- print(cor3$estimate)
corr.df$rho[4] <- print(cor4$estimate)
corr.df$rho[5] <- print(cor5$estimate)
corr.df$rho[6] <- print(cor6$estimate)
corr.df$rho[7] <- print(cor7$estimate)

corr.df$abs <- abs(corr.df$rho)

corr.df.sort <- corr.df[order(corr.df$abs),]
```

# plotting correlations in tornado plot

```{r}

corr.df %>%
  mutate(variable = fct_reorder(variable,abs)) %>%
  ggplot(aes(x=variable, y=rho)) +
  geom_bar(stat="identity", alpha=.6, width=.4) +
  coord_flip() +
  theme_bw() +
  geom_hline(yintercept = 0, size=1) +
  labs(y = "Spearman's rank correlation coefficient", x = expression("N"[2]*"O ratio vs.")) +
  theme(axis.title=element_text(size=14), axis.text=element_text(size=12)) +
  scale_y_continuous(breaks=c(-.5,-.4,-.3,-.2,-.1,0,.1,.2))


```
# Exploring pH effect

Because there were some experiments that manipulated pH and some that didn't, I will separate out those studies and run the correlations again.
-Natural pH studies = 41 with 272 observations
-Adjusted pH studies = 6 with 91 observations
-Missing pH studies = 5 with 52 observations

```{r}
# creating df for each experiment based on if they manipulated pH or not. Include only those with reported pH
natural.pH <- meta.data[c(1:26,50:66,83:92,125:167,187:235,260:269,281:334,346:359,364:401,405:415), ]
adjusted.pH <- meta.data[c(27:46,67:82,93:124,168:176,335:345,402:404), ]

# creating df for studies that are missing pH vs reporting it to rerun correlation and ensure missing data was removed in the previous run
missing.pH <- meta.data[c(47:49,177:186,236:259,270:280,360:363), ]
reported.pH <- meta.data[-c(47:49,177:186,236:259,270:280,360:363), ]

# correlation tests comparing pH manipulations and natural. also quality control of studies with reported pH only to compare to above where rho = -0.4786245 
cor.test(natural.pH$N2O.ratio_trt, natural.pH$Soil.pH, method="spearman")
cor.test(adjusted.pH$N2O.ratio_trt, adjusted.pH$Soil.pH, method="spearman")
cor.test(reported.pH$N2O.ratio_trt, reported.pH$Soil.pH, method="spearman") # same results!
```
- Takeaway: it doesn't really matter if the pH was manipulated or not.
