---
title: "Foltz_Meta-analysis_2021-09-22"
output:
  word_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

# load in necessary packages (make sure you installed them first)
```{r}
library('metafor')
library('tidyverse')
library('dplyr')
library('ggplot2')
```

## Read and prepare data base

```{r}
meta.data <- read.csv("N2O ratio database 2021-09-22.csv", header=TRUE, sep=",")

# clean up dataframe and select relevant columns
meta.data <- meta.data %>% 
  mutate(CN.total = as.numeric(C.N.total), CN.added = as.numeric(C.N.added),
         Soil.texture = as.factor(Soil.texture)) %>% 
  select(1:28)

# make dataframe of samples that have treatment and control
meta.data.wcont.wsd <- meta.data %>%
  filter(!is.na(N2O.ratio_cont), SD_cont > 0, SD_trt > 0)

# Use escalc function to setup meta using mean difference "MD" --> calculates yi and vi
dat <- escalc(measure = "MD", m1i = N2O.ratio_trt, m2i = N2O.ratio_cont,
              sd1i = SD_trt, sd2i = SD_cont, n1i = n, n2i = n,
              data = meta.data.wcont.wsd, append = TRUE)
```

### DerSimonian-Laird "DL" estimate method
### Mixed model without covariates

```{r}
mixmod.0 <- rma(yi, vi, data = dat, method="DL")
mixmod.0
# forest(mixmod.0)
# confint(mixmod.0)
```

## Mixed model (DL method) with covariates

### mixmod.text.pH: texture + pH 

```{r}
mixmod.text.pH <- update(mixmod.0, mods = ~ Soil.texture + Soil.pH)
mixmod.text.pH

mixmod.pH.text <- update(mixmod.0, mods = ~ Soil.pH + Soil.texture)
mixmod.pH.text
```

- these models have the same output regardless of order.

# To get a sense of the importance of each moderator we need to test them individually.

```{r}
mixmod.pH <- update(mixmod.0, mods = ~ Soil.pH)
mixmod.pH


mixmod.text <- update(mixmod.0, mods = ~ Soil.texture)
mixmod.text

anova(mixmod.text,mixmod.0)
```

- It seems that texture is more important than pH. Caveat, the number of observations in each text is different. To test the marginal contribution of the addition of pH or texture we need to remove NAs to have the same observations.

```{r}
mixmod.text2 <- update(mixmod.0, mods = ~ Soil.texture, data = dat[!is.na(dat$Soil.pH),])
mixmod.text2
anova(mixmod.text)

mixmod.02 <- update(mixmod.0, data = dat[!is.na(dat$Soil.pH),])

anova(mixmod.text2,mixmod.02)
anova(mixmod.pH,mixmod.02)
```

```{r}
anova(mixmod.text.pH, mixmod.text2)
anova(mixmod.text.pH, mixmod.pH)
```

- Adding texture to a model with pH improves the fit but not the opposite.

### Models with nutrient covariates added to texture and pH

The next several models add C, N, and CN added

mixmod.1: texture + pH + C
mixmod.2: texture + pH + N
mixmod.3: texture + pH + C + N
mixmod.4: texture + pH + CN
mixmod.5: texture + pH + C + CN
mixmod.6: texture + pH + C + N + CN
mixmod.7: texture + pH + N + CN

```{r}
mixmod.1 <- update(mixmod.0, mods = ~ Soil.texture + Soil.pH + C.added)
mixmod.2 <- update(mixmod.0, mods = ~ Soil.texture + Soil.pH + N.added)
mixmod.3 <- update(mixmod.0, mods = ~ Soil.texture + Soil.pH + C.added + N.added)
mixmod.4 <- update(mixmod.0, mods = ~ Soil.texture + Soil.pH + C.N.added)
mixmod.5 <- update(mixmod.0, mods = ~ Soil.texture + Soil.pH + C.added + C.N.added)
mixmod.6 <- update(mixmod.0, mods = ~ Soil.texture + Soil.pH + C.added + N.added + C.N.added)
mixmod.7 <- update(mixmod.0, mods = ~ Soil.texture + Soil.pH + N.added + C.N.added)
```

```{r}
mixmod.1
#anova(mixmod.1)
mixmod.2
#anova(mixmod.2)
```

Now I will consider C, N, and CN total, which showed up as important in the correlation analysis.

mixmod.1t: texture + pH + total C
mixmod.2t: texture + pH + total N
mixmod.3t: texture + pH + total CN

```{r}
mixmod.1t <- update(mixmod.0, mods = ~ Soil.texture + Soil.pH + Total.C)
mixmod.2t <- update(mixmod.0, mods = ~ Soil.texture + Soil.pH + Total.N)
mixmod.3t <- update(mixmod.0, mods = ~ Soil.texture + Soil.pH + C.N.total)
```

### Comparing models

First, starting with text and pH as reference model.

```{r}
# Addition of C
anova(mixmod.1, mixmod.text.pH)

# Addition of N
anova(mixmod.2, mixmod.text.pH)

# Addition of CN
anova(mixmod.4, mixmod.text.pH)

# Addition of total C
anova(mixmod.1t, mixmod.text.pH)

# Addition of total N
anova(mixmod.2t, mixmod.text.pH)

# Addition of total CN
anova(mixmod.3t, mixmod.text.pH)
```

- Adding C or N improves the fit, but CN does not improve the fit.
- None of the total nutrient variables improved the model fit.

Now I need to run more C or N models with additions of total nutrient variables

mixmod.8: texture + pH + C + total C
mixmod.9: texture + pH + C + total N
mixmod.10: texture + pH + C + total CN
mixmod.11: texture + pH + N + total C
mixmod.12: texture + pH + N + total N
mixmod.13: texture + pH + N + total CN

```{r}
mixmod.8 <- update(mixmod.0, mods = ~ Soil.texture + Soil.pH + C.added + Total.C)
mixmod.9 <- update(mixmod.0, mods = ~ Soil.texture + Soil.pH + C.added + Total.N)
mixmod.10 <- update(mixmod.0, mods = ~ Soil.texture + Soil.pH + C.added + C.N.total)
mixmod.11 <- update(mixmod.0, mods = ~ Soil.texture + Soil.pH + N.added + Total.C)
mixmod.12 <- update(mixmod.0, mods = ~ Soil.texture + Soil.pH + N.added + Total.N)
mixmod.13 <- update(mixmod.0, mods = ~ Soil.texture + Soil.pH + N.added + C.N.total)
```


Next, adding to the C or N model

```{r}
# Set mixmod.text.Ph + C as reduced model, add N
anova(mixmod.3, mixmod.1)

# Set mixmod.text.Ph + C as reduced model, add CN
anova(mixmod.5, mixmod.1)

# Set mixmod.text.Ph + C as reduced model, add total C
anova(mixmod.8, mixmod.1)

# Set mixmod.text.Ph + C as reduced model, add total N
anova(mixmod.9, mixmod.1)

# Set mixmod.text.Ph + C as reduced model, add total CN
anova(mixmod.10, mixmod.1)


# Set mixmod.text.Ph + N as reduced model, add C
anova(mixmod.3, mixmod.2)

# Set mixmod.text.Ph + N as reduced model, add CN
anova(mixmod.7, mixmod.2)

# Set mixmod.text.Ph + N as reduced model, add total C
anova(mixmod.11, mixmod.2)

# Set mixmod.text.Ph + N as reduced model, add total N
anova(mixmod.12, mixmod.2)

# Set mixmod.text.Ph + N as reduced model, add total CN
anova(mixmod.13, mixmod.2)

```

- Adding N or CN on top of texture, pH and C does not significantly improve the fit.
- Adding C or CN on top of texture, pH and N does not significantly improve the fit.

#######################################################################################

## OVERALL CONCLUSION based on above code...
Although the "best" model is mixmod.1 with pH, texture, and added C, the additional improvement from added C or added pH is minimal compared to the dramatic improvement that came from added texture. Therefore, we select the model with texture = mixmod.text

#######################################################################################

# default forest plot in R

```{r}
forest(mixmod.text)
```
- pretty ugly, so improved below

# Forest plots (better than default)

```{r}
mixmod.text

# getting estimates from mixmod.text
tmp <- with(mixmod.text, data.frame(term = rownames(beta), beta = as.vector(beta), ci.lb, ci.ub, se, zval, pval))


# Fixing names (intercept is silty clay loam)
tmp <- tmp %>% 
  mutate(
    term = gsub("Soil.texture", "", term),           
    term = ifelse(term == 'intrcpt', "silty clay loam", term)
  )

# Computing MD and CI
tmp <- tmp %>% 
  mutate(
    int = beta[1],                                   #
    MD = ifelse(term == "silty clay loam", int, beta+int),      #
    CI_LL = MD - 1.96 * se,
    CI_UL = MD + 1.96 * se,
    labs = c("3/1", '4/2', '5/2', '38/5', '10/2', '2/1', '13/1', '3/1')
  )

# getting sample numbers by soil texture
dat %>% 
 group_by(Soil.texture) %>%
 summarise(no_rows = length(Soil.texture))

# Plotting
ggplot(tmp, aes(x = term, y = MD, ymin = CI_LL, ymax = CI_UL)) +
  geom_point(size=3) +
  geom_errorbar(width = 0.15) +
  geom_hline(yintercept = 0) +
  ylim(-1,1) +
  geom_text(aes(y = Inf, label = labs), hjust = 1.2, color = "black") +
  geom_text(x = 8.4, y = Inf, label = "Observations/Studies", hjust = 1.05) +
  coord_flip() +
  ylim(-0.5,1.25) +
  labs(x = "Soil texture", y = "Mean difference") +
  theme_bw() +
  theme(axis.title=element_text(size=13), axis.text=element_text(size=12))

#write.csv(tmp, "tmp.csv")
#write.csv(dat, "dat.csv")

# Bias assessment
ggplot(dat, aes(x = yi, y = vi)) +
  geom_point(size=3) +
  labs(x = "Mean Difference", y = "Variance") +
  theme_bw() +
  theme(axis.title=element_text(size=13), axis.text=element_text(size=12))
cor.test(dat$yi, dat$vi, method="spearman") 
```


# database info

```{r}
meta.data <- read.csv("N2O ratio database 2021-09-22.csv", header=TRUE, sep=",")

# clean up dataframe and select relevant columns
meta.data <- meta.data %>% 
  mutate(CN.total = as.numeric(C.N.total), CN.added = as.numeric(C.N.added),
         Soil.texture = as.factor(Soil.texture)) %>% 
  select(1:28)

# rename first column
colnames(meta.data)[1] <- "author_year"

# make dataframe of samples that have treatment and control
meta.data.wcont.wsd <- meta.data %>%
  filter(!is.na(N2O.ratio_cont), SD_cont > 0, SD_trt > 0)

# Use escalc function to setup meta using mean difference "MD" --> calculates yi and vi
dat <- escalc(measure = "MD", m1i = N2O.ratio_trt, m2i = N2O.ratio_cont,
              sd1i = SD_trt, sd2i = SD_cont, n1i = n, n2i = n,
              data = meta.data.wcont.wsd, append = TRUE)

# getting sample numbers by study
dat %>% 
 group_by(author_year) %>%
 summarise(no_rows = length(author_year))
meta.data %>% 
 group_by(author_year) %>%
 summarise(no_rows = length(author_year))

# getting sample numbers by location
dat %>% 
 group_by(Location) %>%
 summarise(no_rows = length(Location))
meta.data %>% 
 group_by(Location) %>%
 summarise(no_rows = length(Location))

# getting sample numbers by substrate source
dat %>% 
 group_by(Substrate.source) %>%
 summarise(no_rows = length(Substrate.source))
meta.data %>% 
 group_by(Substrate.source) %>%
 summarise(no_rows = length(Substrate.source))

# getting sample numbers by soil texture
dat %>% 
 group_by(Soil.texture) %>%
 summarise(no_rows = length(Soil.texture))
meta.data %>% 
 group_by(Soil.texture) %>%
 summarise(no_rows = length(Soil.texture))

# getting sample numbers by added C type
dat %>% 
 group_by(Added.C.type) %>%
 summarise(no_rows = length(Added.C.type))
meta.data %>% 
 group_by(Added.C.type) %>%
 summarise(no_rows = length(Added.C.type))

# getting sample numbers by added N type
dat %>% 
 group_by(Added.N.type) %>%
 summarise(no_rows = length(Added.N.type))
meta.data %>% 
 group_by(Added.N.type) %>%
 summarise(no_rows = length(Added.N.type))

# getting sample numbers by headspace gas
dat %>% 
 group_by(Headspace.gas) %>%
 summarise(no_rows = length(Headspace.gas))
meta.data %>% 
 group_by(Headspace.gas) %>%
 summarise(no_rows = length(Headspace.gas))

# getting sample numbers by inhibitors
dat %>% 
 group_by(Inhibitors) %>%
 summarise(no_rows = length(Inhibitors))
meta.data %>% 
 group_by(Inhibitors) %>%
 summarise(no_rows = length(Inhibitors))

# summary for pH
summary(meta.data$Soil.pH)
summary(dat$Soil.pH)

# summary for BD
summary(meta.data$Soil.BD)
summary(dat$Soil.BD)

# summary for substrate amount
summary(meta.data$Substrate.amt)
summary(dat$Substrate.amt)

# summary for incubation temp
summary(meta.data$Temperature)
summary(dat$Temperature)

# summary for incubation time
summary(meta.data$Incubation.time)
summary(dat$Incubation.time)

# summary for added C
summary(meta.data$C.added)
summary(dat$C.added)

# summary for initial C
summary(meta.data$C.initial)
summary(dat$C.initial)

# summary for added N
summary(meta.data$N.added)
summary(dat$N.added)

# summary for initial N
summary(meta.data$N.initial)
summary(dat$N.initial)

```